---
title: "APM1110 FA5 Cellona, Bryan Dion"
subtitle: '**Github link: https://github.com/bryandioncellona/APM1110/blob/main/FA5/APM1110_FA5_Cellona_BryanDion.Rmd**'
date: "September 17, 2026"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
fmt <- function(x) sprintf("%.4f", x)
```

# 1. Discrete Random Variables

A discrete random variable $X$ has the probability mass function

$$
P(X=x)=\begin{cases}
k(2x+1), & x=1,2,3,4,\\
0, & \text{otherwise}.
\end{cases}
$$

## (a) Find the constant $k$.

Since probabilities must sum to 1,

$$
\begin{aligned}
\sum_{x=1}^{4}k(2x+1)&=1,\\
k(3+5+7+9)&=1,\\
24k&=1,\\
k&=\frac{1}{24}\approx\boxed{0.0417}.
\end{aligned}
$$

```{r}
x <- 1:4
k <- 1 / sum(2 * x + 1)
px <- k * (2 * x + 1)
cat("k =", fmt(k))
```

\newpage

## (b) Compute the expected value $E[X]$ and the variance $\operatorname{Var}(X)$.

First compute the expected value:

$$
\begin{aligned}
E[X]&=\sum_{x=1}^{4}xP(X=x)\\
&=\frac{1(3)+2(5)+3(7)+4(9)}{24}\\
&=\frac{70}{24}=\frac{35}{12}\approx\boxed{2.9167}.
\end{aligned}
$$

Next compute the second moment:

$$
\begin{aligned}
E[X^2]&=\sum_{x=1}^{4}x^2P(X=x)\\
&=\frac{1^2(3)+2^2(5)+3^2(7)+4^2(9)}{24}\\
&=\frac{3+20+63+144}{24}=\frac{115}{12}\approx\boxed{9.5833}.
\end{aligned}
$$

Therefore,

$$
\begin{aligned}
\operatorname{Var}(X)&=E[X^2]-(E[X])^2\\
&=\frac{115}{12}-\left(\frac{35}{12}\right)^2\\
&=\frac{1380-1225}{144}=\frac{155}{144}
\approx\boxed{1.0764}.
\end{aligned}
$$

```{r}
EX <- sum(x * px)
EX2 <- sum(x^2 * px)
var_X <- EX2 - EX^2
cat("E[X] =", fmt(EX))
cat("E[X^2] =", fmt(EX2))
cat("Var(X) =", fmt(var_X))
```

## (c) Calculate $P(X\geq 3)$.

$$
P(X\geq3)=P(X=3)+P(X=4)
=\frac{7}{24}+\frac{9}{24}=\frac{2}{3}
\approx\boxed{0.6667}.
$$

```{r}
p_X_ge3 <- sum(px[x >= 3])
cat("P(X >= 3) =", fmt(p_X_ge3))
```

# 2. Binomial Distribution

An IT system detects that 8% of incoming external requests are malicious. A random sample contains 20 requests. Let $B$ be the number of malicious requests. Assuming independent requests and a constant malicious-request probability,

$$B\sim\operatorname{Binomial}(n=20,p=0.08),\qquad 1-p=0.92.$$

The probability formula is

$$P(B=b)=\binom{n}{b}p^b(1-p)^{n-b},\qquad
\binom{n}{b}=\frac{n!}{b!(n-b)!}.$$

```{r}
n <- 20
p <- 0.08
```

## (a) What is the probability that exactly 2 requests are malicious?

$$
\begin{aligned}
P(B=2) &= \binom{20}{2}(0.08)^2(0.92)^{18}\\
&=\frac{20(19)}{2}(0.0064)(0.92)^{18}\\
&=190(0.0064)(0.92)^{18}\\
&\approx0.2710906052\approx\boxed{0.2711}.
\end{aligned}
$$

```{r}
p_B_eq2 <- dbinom(2, size = n, prob = p)
cat("P(B = 2) =", fmt(p_B_eq2))
```

## (b) What is the probability that at least 3 requests are malicious?

Use the complement of 0, 1, or 2 malicious requests:

$$P(B\geq3)=1-[P(B=0)+P(B=1)+P(B=2)].$$

The component probabilities are

$$
\begin{aligned}
P(B=0)&=(0.92)^{20}\approx0.1886933292,\\
P(B=1)&=20(0.08)(0.92)^{19}\approx0.3281623116,\\
P(B=2)&\approx0.2710906052.
\end{aligned}
$$

Thus,

$$
\begin{aligned}
P(B\geq3)&\approx1-(0.1886933292+0.3281623116+0.2710906052)\\
&\approx0.2120537540\approx\boxed{0.2121}.
\end{aligned}
$$

```{r}
# lower.tail = FALSE returns P(B > 2), which equals P(B >= 3).
p_B_ge3 <- pbinom(2, size = n, prob = p, lower.tail = FALSE)
cat("P(B >= 3) =", fmt(p_B_ge3))
```

## (c) Determine the mean and standard deviation of the number of malicious requests.

$$
\begin{aligned}
\mu_B&=np=20(0.08)=\boxed{1.6000},\\
\sigma_B^2&=np(1-p)=20(0.08)(0.92)=1.472,\\
\sigma_B&=\sqrt{np(1-p)}=\sqrt{1.472}
\approx\boxed{1.2133}.
\end{aligned}
$$

Both the mean and standard deviation are expressed in malicious requests.

```{r}
mean_B <- n * p
sd_B <- sqrt(n * p * (1 - p))
cat("Mean =", fmt(mean_B), "malicious requests")
cat("Standard deviation =", fmt(sd_B), "malicious requests")
```

# 3. Poisson Distribution

A server receives an average of 4.5 web traffic hits per minute, with arrivals following a Poisson process. For an interval of $t$ minutes, the expected number of hits is $\lambda=4.5t$.

For a Poisson random variable $H$,

$$P(H=h)=\frac{e^{-\lambda}\lambda^h}{h!},\qquad h=0,1,2,\ldots.$$

```{r}
rate <- 4.5
lambda_1 <- rate * 1
```

## (a) Find the probability of receiving exactly 6 hits in a given minute.

Here $\lambda=4.5$:

$$
\begin{aligned}
P(H=6)&=\frac{e^{-4.5}(4.5)^6}{6!}\\
&=\frac{e^{-4.5}(8303.765625)}{720}\\
&\approx0.1281201439\approx\boxed{0.1281}.
\end{aligned}
$$

```{r}
p_H_eq6 <- dpois(6, lambda = lambda_1)
cat("P(H = 6) =", fmt(p_H_eq6))
```

## (b) Find the probability of receiving at most 2 hits in a given minute.

“At most 2” includes 0, 1, and 2 hits:

$$
\begin{aligned}
P(H\leq2)&=\sum_{h=0}^{2}\frac{e^{-4.5}(4.5)^h}{h!}\\
&=e^{-4.5}\left(1+4.5+\frac{4.5^2}{2}\right)\\
&=e^{-4.5}(1+4.5+10.125)\\
&=e^{-4.5}(15.625)\\
&\approx0.1735780709\approx\boxed{0.1736}.
\end{aligned}
$$

```{r}
p_H_le2 <- ppois(2, lambda = lambda_1)
cat("P(H <= 2) =", fmt(p_H_le2))
```

## (c) Calculate the probability of receiving more than 10 hits in a 2-minute interval.

Let $Y$ count hits in two minutes. Adjust the parameter for the time interval:

$$\lambda_2=4.5(2)=9,\qquad Y\sim\operatorname{Poisson}(9).$$

“More than 10” means 11 or more. Use the complement:

$$
\begin{aligned}
P(Y>10)&=1-P(Y\leq10)\\
&=1-e^{-9}\sum_{y=0}^{10}\frac{9^y}{y!}\\
&=1-e^{-9}\left(1+9+\frac{9^2}{2!}+\cdots+\frac{9^{10}}{10!}\right)\\
&\approx1-0.7059883203\\
&\approx0.2940116797\approx\boxed{0.2940}.
\end{aligned}
$$

```{r}
lambda_2 <- rate * 2
# The upper tail excludes 10, as required by "more than 10".
p_Y_gt10 <- ppois(10, lambda = lambda_2, lower.tail = FALSE)
cat("Two-minute mean =", fmt(lambda_2), "hits")
cat("P(Y <= 10) =", fmt(ppois(10, lambda_2)))
cat("P(Y > 10) =", fmt(p_Y_gt10))
```

\newpage

# 4. Continuous Probability and Normal Distribution

The processing time $T$ of a complex database query is normally distributed with mean $\mu=120$ milliseconds and standard deviation $\sigma=15$ milliseconds:

$$T\sim N(120,15^2),\qquad Z=\frac{T-\mu}{\sigma}.$$

Write $\Phi(z)=P(Z\leq z)$ for the standard normal cumulative distribution function. Do not round the standardized values before calculating probabilities.

```{r}
mu <- 120
sigma <- 15
```

## (a) What proportion of queries take longer than 135 milliseconds to process?

$$z=\frac{135-120}{15}=1.$$

Therefore,

$$
\begin{aligned}
P(T>135)&=P(Z>1)=1-\Phi(1)\\
&\approx1-0.8413447461\\
&\approx0.1586552539\approx\boxed{0.1587}.
\end{aligned}
$$

This corresponds to approximately **15.8655%** of queries.

```{r}
p_T_gt135 <- pnorm(135, mean = mu, sd = sigma, lower.tail = FALSE)
cat("P(T > 135) =", fmt(p_T_gt135))
cat("Percentage =", fmt(100 * p_T_gt135), "%")
```

\newpage

## (b) Find the probability that a randomly chosen query takes between 100 and 130 milliseconds

Standardize both endpoints:

$$z_L=\frac{100-120}{15}=-\frac43,\qquad
z_U=\frac{130-120}{15}=\frac23.$$

Then

$$
\begin{aligned}
P(100<T<130)&=P\left(-\frac43<Z<\frac23\right)\\
&=\Phi\left(\frac23\right)-\Phi\left(-\frac43\right)\\
&\approx0.7475074625-0.0912112197\\
&\approx0.6562962427\approx\boxed{0.6563}.
\end{aligned}
$$

Since $T$ is continuous, including the endpoints does not change the probability.

```{r}
z_lower <- (100 - mu) / sigma
z_upper <- (130 - mu) / sigma
p_T_between <- pnorm(130, mu, sigma) - pnorm(100, mu, sigma)
cat("Lower z =", fmt(z_lower))
cat("Upper z =", fmt(z_upper))
cat("P(100 < T < 130) =", fmt(p_T_between))
```

## (c) Find the 95th percentile of query processing times.

Let $t_{0.95}$ satisfy $P(T\leq t_{0.95})=0.95$. The standard normal 95th percentile is

$$z_{0.95}=\Phi^{-1}(0.95)\approx1.6448536270.$$

Solve for the processing time:

$$
\begin{aligned}
\frac{t_{0.95}-120}{15}&=z_{0.95},\\
t_{0.95}&=\mu+\sigma z_{0.95}\\
&=120+15(1.6448536270)\\
&\approx\boxed{144.6728\text{ milliseconds}}.
\end{aligned}
$$

Approximately 95% of queries take no more than 144.6728 milliseconds; approximately 5% take longer.

```{r}
z_95 <- qnorm(0.95)
t_95 <- qnorm(0.95, mean = mu, sd = sigma)
cat("Standard normal 95th percentile =", fmt(z_95))
cat("95th percentile =", fmt(t_95), "milliseconds")
```
