---
title: "SEC 2-FA2 GROUP 3-CELLONA, BD; RICAFORT, NF-FA2"
output: pdf_document
date: "September 1, 2026"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## 2. An experiment consists of tossing two fair coins. Use R to simulate this experiment 100 times and obtain the relative frequency of each possible outcome. Hence, estimate the probability of getting one head and one tail in any order.

```{r}
# Toss two coins 100 times
coin1 <- sample(c("H", "T"), 100, replace = TRUE)
coin2 <- sample(c("H", "T"), 100, replace = TRUE)

# Combine the outcomes
outcomes <- paste(coin1, coin2, sep = "")

# Frequency of each outcome
table(outcomes)
```

## Relative Frequencies

```{r}
# Find the relative frequency
table(outcomes) / 100
```

## Probability of One Head and One Tail

One head and one tail can occur as **HT** or **TH**.

```{r}
# Estimate the probability
probability <- sum(outcomes == "HT" | outcomes == "TH") / 100

probability
```

The estimated probability of getting one head and one tail is given by the simulation result above. It should be close to **0.5**.