---
title: "SEC 2-FA4 GROUP 3-CELLONA, BD; RICAFORT, NF-FA4"
subtitle: "**Github link: https://github.com/bryandioncellona/APM1110/blob/main/FA4/SEC%202-FA4%20GROUP%203-CELLONA%2C%20BD%3B%20RICAFORT%2C%20NF-FA4.Rmd**"
output: pdf_document
date: "September 9, 2026"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2. A binary communication channel carries data as one of two sets of signals denoted by 0 and 1. Owing to noise, a transmitted 0 is sometimes received as a 1, and a transmitted 1 is sometimes received as a 0. For a given channel, it can be assumed that a transmitted 0 is correctly received with probability 0.95, and a transmitted 1 is correctly received with probability 0.75. Also, 70% of all messages are transmitted as a 0. If a signal is sent, determine the probability that:

```{r}
binary <- data.frame(
  Correct = c(0.95, 0.75),
  Error   = c(0.05, 0.25),
  Sent    = c(0.70, 0.30),
  row.names = c("0", "1")
)

binary
```

### (a) a 1 was received;

$P(R = 1) = P(R = 1 | T = 0) \cdot P(T = 0) + P(R = 1 | T = 1) \cdot P(T = 1)$

```{r}
prob_get1 <- (binary["0", "Error"] * binary["0", "Sent"]) +
             (binary["1", "Correct"] * binary["1", "Sent"])

cat(round(prob_get1 * 100, 2), "%", sep = "")
```

### (b) a 1 was transmitted given than a 1 was received.

$P(T = 1 | R = 1) = \frac{P(T = 1) \cdot P(R = 1 | T = 1)}{P(R = 1)}$

```{r}
prob_send1_get1 <- (binary["1", "Sent"] * binary["1", "Correct"]) /
                   prob_get1

cat(round(prob_send1_get1 * 100, 2), "%", sep = "")
```

## 7. There are three employees working at an IT company: Jane, Amy, and Ava, doing 10%, 30%, and 60% of the programming, respectively. 8% of Jane’s work, 5% of Amy’s work, and just 1% of Ava‘s work is in error. What is the overall percentage of error? If a program is found with an error, who is the most likely person to have written it?

```{r}
# Given probabilities
P_Jane <- 0.10
P_Amy  <- 0.30
P_Ava  <- 0.60

P_E_Jane <- 0.08
P_E_Amy  <- 0.05
P_E_Ava  <- 0.01


# Overall percentage of error
P_Error <- (P_Jane * P_E_Jane) +
  (P_Amy * P_E_Amy) +
  (P_Ava * P_E_Ava)

cat("Overall Error Probability:", P_Error, "\n")
cat("Overall Error Percentage:", P_Error * 100, "%\n\n")


# If a program is found with an error,
# who is the most likely person to have written it?
employee <- c("Jane", "Amy", "Ava")

# Bayes' Theorem
P_Jane_given_Error <- (P_Jane * P_E_Jane) / P_Error
P_Amy_given_Error  <- (P_Amy * P_E_Amy) / P_Error
P_Ava_given_Error  <- (P_Ava * P_E_Ava) / P_Error

P_given_Error <- c(
  P_Jane_given_Error,
  P_Amy_given_Error,
  P_Ava_given_Error
)

# Create a table of results
E_table <- data.frame(
  Employee = employee,
  Probability = round(P_given_Error, 4),
  Percentage = round(P_given_Error * 100, 2)
)

print(E_table)

# Find the employee with the highest probability
most_likely <- which.max(P_given_Error)
most_likely_employee <- employee[most_likely]
most_likely_prob <- P_given_Error[most_likely]

cat("Most likely employee to have written the program:",
    most_likely_employee, "\n")

cat("Probability:",
    round(most_likely_prob, 4), "\n")

cat("Percentage:",
    round(most_likely_prob * 100, 2), "%\n")
```
