---
title: "APM1110 FA6 Cellona, Bryan Dion"
subtitle: '**Github link: https://github.com/bryandioncellona/APM1110/blob/main/FA6/APM1110-FA6.Rmd**'
date: "September 19, 2026"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  comment = "##", fig.width = 8, fig.height = 5, dpi = 150
)
options(width = 80)
```

# I. Geometric Distribution

Provide an R code for the geometric distribution. The geometric distribution is a probability distribution that models the number of trials required to achieve the first success in a sequence of Bernoulli trials, where each trial has a constant probability of success.

## 1. Set the probability of success

```{r}
# Each trial has a 20% probability of success.
p <- 0.2
```

## 2. Generate 1000 geometric random variables

```{r}
# Fix the random seed and RNG settings so the results can be reproduced.
RNGkind(kind = "Mersenne-Twister", normal.kind = "Inversion",
        sample.kind = "Rejection")
set.seed(1110)
# rgeom() counts failures; add 1 to include the successful trial.
x <- rgeom(n = 1000, prob = p) + 1
```

There are 1000 simulated observations. Each observation represents one sequence of trials ending at its first success; it does not represent a single Bernoulli trial.

## 3. Calculate basic statistics

```{r}
# Calculate the sample mean, sample variance, and sample standard deviation.
# var() uses the sample denominator n - 1.
mean_x <- mean(x)
var_x <- var(x)
sd_x <- sd(x)
```

\newpage

## 4. Print the required results

The first output lists all 1000 simulated trial counts. `sprintf()` formats each summary statistic with exactly two digits after the decimal point.

```{r}
cat("Number of trials required to achieve first success:")
print(x)
cat(sprintf("Mean (in 2 decimal places): %.2f", mean_x))
cat(sprintf("Variance (in 2 decimal places): %.2f", var_x))
cat(sprintf("Standard deviation (in 2 decimal places): %.2f", sd_x))
```

## 5. Plot the histogram

```{r}
# Half-integer boundaries put one integer trial count in each bin.
hist(
  x,
  breaks = seq(0.5, max(x) + 0.5, by = 1),
  freq = TRUE,
  col = "#69A9C7",
  border = "white",
  main = "Trials Required to Achieve the First Success",
  xlab = "Number of trials (including the successful trial)",
  ylab = "Frequency"
)
```

The distribution is right-skewed. Short waits are more common while some sequences require many trials before the first success.

# II. Hypergeometric Distribution

Consider a plant manufacturing IC chips of which 10% are expected to be defective. The chips are packed in boxes for export. Before transportation, a sample is drawn from each box. Estimate the probability that the sample contains more than 10% defectives, when:

## 1. A sample of 10 from a box of 40

Here, $N=40$, $K=4$, $N-K=36$, and $s=10$. Therefore,

$$
P(D>1)=1-
\frac{\binom{4}{0}\binom{36}{10}
+\binom{4}{1}\binom{36}{9}}{\binom{40}{10}}.
$$

```{r}
box_40 <- 40
sample_size <- 10
defective_40 <- 4
good_40 <- box_40 - defective_40

# More than 10% defective in a sample of 10 means more than 1 defective.
prob_40 <- phyper(q = 1, m = defective_40, n = good_40,
                 k = sample_size, lower.tail = FALSE)

# Independently evaluate the complement using the probability mass function.
p0_40 <- dhyper(0, m = defective_40, n = good_40, k = sample_size)
p1_40 <- dhyper(1, m = defective_40, n = good_40, k = sample_size)
check_40 <- 1 - p0_40 - p1_40
stopifnot(isTRUE(all.equal(prob_40, check_40)))

cat(sprintf("P(D = 0), box of 40: %.2f", p0_40))
cat(sprintf("P(D = 1), box of 40: %.2f", p1_40))
cat(sprintf("P(D > 1), box of 40: %.2f", prob_40))
cat(sprintf("Probability as a percentage: %.2f%%", 100 * prob_40))
```

The probability is approximately **`r sprintf("%.2f", prob_40)`** or **`r sprintf("%.2f", 100 * prob_40)`%**. This is the chance that a random sample of 10 chips contains at least 2 of the box's 4 defective chips.

## 2. A sample of 10 from a box of 5000

Here, $N=5000$, $K=500$, $N-K=4500$, and $s=10$. Therefore,

$$
P(D>1)=1-
\frac{\binom{500}{0}\binom{4500}{10}
+\binom{500}{1}\binom{4500}{9}}{\binom{5000}{10}}.
$$

```{r}
box_5000 <- 5000
defective_5000 <- 500
good_5000 <- box_5000 - defective_5000

prob_5000 <- phyper(q = 1, m = defective_5000, n = good_5000,
                   k = sample_size, lower.tail = FALSE)

# Check the upper-tail result by calculating 1 - P(D = 0) - P(D = 1).
p0_5000 <- dhyper(0, m = defective_5000, n = good_5000, k = sample_size)
p1_5000 <- dhyper(1, m = defective_5000, n = good_5000, k = sample_size)
check_5000 <- 1 - p0_5000 - p1_5000
stopifnot(isTRUE(all.equal(prob_5000, check_5000)))

cat(sprintf("P(D = 0), box of 5000: %.2f", p0_5000))
cat(sprintf("P(D = 1), box of 5000: %.2f", p1_5000))
cat(sprintf("P(D > 1), box of 5000: %.2f", prob_5000))
cat(sprintf("Probability as a percentage: %.2f%%", 100 * prob_5000))
```

The probability is approximately **`r sprintf("%.2f", prob_5000)`** or **`r sprintf("%.2f", 100 * prob_5000)`%**. This is the chance that a random sample of 10 chips contains at least 2 of the box's 500 defective chips.
